---
title: "National Waiting List"
format:
    dashboard:
        nav-buttons:
          - icon: github
            href: https://github.com/The-Strategy-Unit/nhs-waiting-list-explorer
---



# National Table {.page}

## Row

This is the explorer for the national waiting list. You can use it by...

Expand the sidebar on the right to read more about the key metrics.

## Row
```{r}
finalized_table <- readRDS("data/finalized_table.rds")

finalized_table |>
dplyr::filter(Treatment_Function_Code == "C_999") |>
dplyr::select(Provider_Name, Queue_Ratio, Percentile_Pressure, Load) |>
reactable::reactable()
```

# Some info {.sidebar}

Here is some extra info. It appears on _every page_

You could explain the key metrics here if you want

# Detailed Table {.page}

```{r}
source("R/display_detailed_national_table.R")
tbl_widget
```

# Key metrics {.page}

## Row

Space to explain the key metrics

- **Percentile Pressure**: 
- **Queue Ratio**:
- **Load**:

# About {.page}

## Row

::: {.card}

### HACA 2025 presentation

{{< video https://www.youtube.com/watch?v=OIJ4C-8qEi8 >}}
:::

::: {.card}
### Paper

Richard M Wood, Sebastian S Fox, Lucy E Morgan, Neil S Walton,
The NHS waiting list in England must halve to reach waiting time targets,
[Future Healthcare Journal, Volume 12, Issue 4, 2025](https://www.sciencedirect.com/science/article/pii/S2514664525002577)
:::



## Row

::: {.card}

### Contact

For more information about this tool please contact ...
:::
