---
title: "National Waiting List Explorer"
format:
    dashboard:
        nav-buttons:
          - icon: github
            href: https://github.com/The-Strategy-Unit/nhs-waiting-list-explorer
---



# National Table {.page}

## Row
We rank each provider nationally according to their waiting list performance. 
We assess the size, distribution and improvement of each provider trust respectively with the 
queue size ratio, pressure, and load (decribed further in the key metrics tab). 
A pressure of one or lower meets the national waiting list standard.




## Row
```{r}
finalized_table <- readRDS("data/finalized_table.rds")

national_view <- finalized_table |>
  dplyr::filter(Treatment_Function_Code == "C_999") |>
  dplyr::select(Provider_Name, Queue_Ratio, Percentile_Pressure, Load)

# Calculate means and standard deviations
queue_ratio_mean <- mean(national_view$Queue_Ratio, na.rm = TRUE)
queue_ratio_sd <- sd(national_view$Queue_Ratio, na.rm = TRUE)
percentile_pressure_mean <- mean(national_view$Percentile_Pressure, na.rm = TRUE)
percentile_pressure_sd <- sd(national_view$Percentile_Pressure, na.rm = TRUE)
load_mean <- mean(national_view$Load, na.rm = TRUE)
load_sd <- sd(national_view$Load, na.rm = TRUE)

# Calculate Z-scores and composite score
national_view <- national_view |>
  dplyr::mutate(
    Queue_Ratio_Z = (Queue_Ratio - queue_ratio_mean) / queue_ratio_sd,
    Percentile_Pressure_Z = (Percentile_Pressure - percentile_pressure_mean) / percentile_pressure_sd,
    Load_Z = (Load - load_mean) / load_sd,
    Score = (Queue_Ratio_Z + Percentile_Pressure_Z + Load_Z) / 3,
    Ranking = rank(Score, ties.method = "first")
  ) |>
  dplyr::arrange(Ranking) |>
  dplyr::select(Provider_Name, Queue_Ratio, Percentile_Pressure, Load, Score, Ranking)
  reactable::reactable(
    national_view,
    defaultPageSize = 10,
    filterable = TRUE,
    columns = list(
      Provider_Name = reactable::colDef(name = "Provider", filterable = TRUE),
      Queue_Ratio = reactable::colDef(name = "Queue Size Ratio", filterable = FALSE),
      Percentile_Pressure = reactable::colDef(name = "Pressure", filterable = FALSE),
      Load = reactable::colDef(filterable = FALSE),
      Ranking = reactable::colDef(width = 100, filterable = FALSE),
      Score = reactable::colDef(name = "Overall Score", format = reactable::colFormat(digits = 2), filterable = FALSE)
    )
  )
```

# Some info {.sidebar}

```{r}
#| echo: false
#| include: false
#| eval: false
#| message: false
#| warning: false

if (knitr::opts_knit$get('rmarkdown.pandoc.to') == 'html') {
  if (knitr::opts_current$get('tab') == 'National Table') {
    cat("We consider three key statistics: 

1. **The queue size ratio** which judges the current list size relative to what the list should be given the demand observed over the last year. The queue ration measures the size of the list.

2. **The Pressure** which is the ratio between the 92nd percentile of wait time on the list and 18 weeks. Pressure of 1 or lower meets the NHS target of 92% having patients waited 18 weeks.  Pressure above 1 does not. Pressure measures the shape of the list.

3. **The Load** is the ratio between observed rate of arrivals and departures. I.e. for every departure we make how many arrivals should we expect. This is typically close to 1. A value above one means the list has increased below 1 the list has decreased. Thus is a measurement of change and improvement.

The score is the average z-score (value minus mean over standard deviation) of the three statistics collected. These are then ranked to give a ranking of the providers.")
  } else if (knitr::opts_current$get('tab') == 'Detailed Table') {
    cat("Content for Detailed Table goes here.")
  }
}
```



# Specialties Table {.page}

## Row
This view ranks specialties by key pressure metrics using precomputed scores.

## Row
```{r}
specialties_ranked <- readRDS("data/specialties_ranked.rds")

reactable::reactable(
  specialties_ranked,
  defaultPageSize = 10,
  filterable = TRUE,
  columns = list(
    Treatment_Function_Code = reactable::colDef(name = "Specialty Code", filterable = TRUE),
    Treatment_Function = reactable::colDef(name = "Specialty", filterable = TRUE),
    Queue_Ratio = reactable::colDef(name = "Queue Size Ratio"),
    Percentile_Pressure = reactable::colDef(name = "Pressure"),
    Load = reactable::colDef(name = "Load"),
    Score = reactable::colDef(name = "Overall Score", format = reactable::colFormat(digits = 2)),
    Ranking = reactable::colDef(name = "Rank", width = 100)
  )
)
```


# Detailed Table {.page}

```{r}
# Load required data and define max_report_date
finalized_table <- readRDS("data/finalized_table.rds")

source("R/display_detailed_national_table.R")
tbl_widget
```

# Key metrics {.page}

## Row

Space to explain the key metrics

- **Percentile Pressure**: 
- **Queue Ratio**:
- **Load**:

# About {.page}

## Row

::: {.card}

### HACA 2025 presentation

{{< video https://www.youtube.com/watch?v=OIJ4C-8qEi8 >}}
:::

::: {.card}
### Paper

Richard M Wood, Sebastian S Fox, Lucy E Morgan, Neil S Walton,
The NHS waiting list in England must halve to reach waiting time targets,
[Future Healthcare Journal, Volume 12, Issue 4, 2025](https://www.sciencedirect.com/science/article/pii/S2514664525002577)
:::



## Row

::: {.card}

### Contact

For more information about this tool please contact ...
:::
